---
title: "Summer Juvenile Fish"
author: "Kevin See^[Biomark Inc.]"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    toc: yes
    toc_depth: 1
vignette: >
  %\VignetteIndexEntry{Summer_Juveniles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
csl: /Users/kevin/Dropbox/Bibliography/StyleFiles/ecology.csl
bibliography:
- /Users/kevin/Dropbox/Bibliography/Research.bib
- /Users/kevin/Dropbox/Bibliography/SoftwareCitations.bib
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

library(pander)
# options for table formatting
panderOptions('big.mark', ',')
# panderOptions('digits', 3)
# panderOptions('round', 3)
panderOptions('keep.trailing.zeros', FALSE)
panderOptions('table.split.table', Inf)

library(captioner)
tab_nums = captioner(prefix = 'Table')
fig_nums = captioner()


# load packages for analysis
library(QRFcapacity)
library(tidyverse)
library(janitor)
library(magrittr)
library(minerva)
library(sf)
library(quantregForest)
library(survey)
library(MuMIn)

# set default theme for ggplot
theme_set(theme_bw())
```

```{r}
# determine which set of fish/habitat data to use
data("fh_sum_dash_2014_17")
fish_hab = fh_sum_dash_2014_17 %>%
  mutate_at(vars(Watershed, Year),
            list(as.factor))

# data("fh_sum_champ_2017")
# fish_hab = fh_sum_champ_2017 %>%
#   mutate_at(vars(Watershed, Year),
#             list(as.factor))

# and the appropriate habitat dictionrary to go with it
data("hab_dict_2017")
hab_dict = hab_dict_2017

# all the related habitat data
data('champ_dash')
hab_data = champ_dash

data('champ_dash_avg')
hab_avg = champ_dash_avg

# data("champ_site_2011_17")
# hab_data = champ_site_2011_17
# 
# data("champ_site_2011_17_avg")
# hab_avg = champ_site_2011_17_avg

# add temperature metrics
data("champ_temps")
hab_avg %<>%
  left_join(champ_temps %>%
              as_tibble() %>%
              select(Site, avg_aug_temp = S2_02_11) %>%
              distinct())


```


# Introduction

# Data

The fish data was collected by various agencies during the summer low-flow season at many of the same sites surveyed using the CHaMP protocol. Survey methods included mark-recapture, three-pass removal sampling, two-pass removal sampling, and single-pass electrofishing, as well as snorkeling. We estimated juvenile fish abundance (and density) at all sites where fish survey data were available. Three-pass removal estimates used the Carle-Strub estimator (@Carle1978), following advice from @Hedger2013. Two-pass removal estimates used the estimator described by @Seber2002. Mark-recapture estimates used Chapman’s modified Lincoln-Peterson estimator (@Chapman1951) and were deemed valid if they met the criteria described in @Robson1964. These estimates were made using the `removal` and `mrClosed` functions from the FSA package (@Ogle2017) in R software (@Rsoftware2015). Snorkel counts were transformed to abundance estimates using paired snorkel-electrofishing sites to calibrate snorkel counts. 

Many sites were sampled in multiple years. To avoid any chance of pseudo-replication (since habitat data is not usually expected to vary much year-to-year), for any site with multiple surveys, we chose the survey with the highest estimated density. We used this criteria because we are interested in estimating carrying capacity, so using the highest observed density seemed appropriate.

***All sites fell within the range of steelhead, but many of them were outside the range of spring/summer Chinook salmon. However, defining the range of sp/su Chinook is tricky, and it is unclear if the reported data for Chinook salmon includes only sites considered with the range of Chinook, or if we should exclude data from some sites based on our best understanding of the species' range. Currently, we are excluding sites outside the Chinook domain. We created our own list of sites in the John Day that were within a Chinook domain, based on which sites Chinook were found at some point (and downstream of those). ***

```{r, eval = T}
library(maptools)
data("chnk_domain")

# which sites were sampled for Chinook? 
chnk_samps = fish_hab %>%
  filter(Species == 'Chinook') %>%
  select(Site:Lon, N) %>%
  distinct() %>%
  st_as_sf(coords = c('Lon', 'Lat'),
           crs = 4326) %>%
  st_transform(st_crs(chnk_domain))

# set snap distance (in meters)
st_crs(chnk_samps)
snap_dist = 1000

# which of those sites are in Chinook domain?
chnk_sites = chnk_samps %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  select(-nearest_line_id, -snap_dist) %>%
  # include some sites in the John Day where Chinook were found (or seemed to be close to sites where Chinook were found)
  rbind(st_read('data/raw/domain/Chnk_JohnDay_TrueObs.shp') %>%
          st_transform(st_crs(chnk_domain)) %>%
          select(-in_range))

# ggplot() +
#   geom_sf(data = chnk_samps,
#           aes(color = 'Sampled')) +
#   geom_sf(data = chnk_sites,
#           aes(color = 'In Chnk Range')) +
#   theme(axis.text = element_blank())

fish_hab %<>%
  filter(Species == 'Steelhead' |
           (Species == 'Chinook' & Site %in% chnk_sites$Site))
```

After matching up the fish abundances with sites where we have habitat data, we are left with the following number of sites.

```{r}
fish_hab %>%
  xtabs(~ Watershed + Species + Year, .) %>%
  addmargins(margin = c(1, 3)) %>%
  pander()
```

```{r}
fish_hab %>%
  tabyl(Watershed, Year, Species) %>%
  adorn_totals(c('row', 'col')) %>%
  adorn_percentages('col') %>%
  adorn_pct_formatting(digits = 0,
                       rounding = 'half up') %>%
  adorn_ns %>%
  adorn_title %>%
  knitr::kable()

```

# Selecting Covariates

A key step in developing a QRF model to predict fish capacities is selecting the habitat covariates to include in the model. Random forest models naturally incorporate interactions between correlated covariates, which is essential since nearly all habitat variables are considered correlated to one degree or another. However, we aimed to avoid overly redundant variables (i.e., variables that measure similar aspects of the habitat). Further, including too many covariates can result in overfitting of the model (e.g., including as many covariates as data points).

## Methods

We used the Maximal Information-Based Nonparametric Exploration (MINE) class of statistics (@Reshef2011) to determine those habitat characteristics (covariates) most highly associated with observed parr densities. We calculated the maximal information coefficient (MIC), using the R package `minerva` (@Albanese2013), to measure the strength of the linear or non-linear association between two variables (@Reshef2011). The MIC value between each of the measured habitat characteristics and the response variable, juvenile fish density (fish/m), was used to inform decisions on which habitat covariates to include in the QRF parr capacity model.
	Habitat metrics were first grouped into broad categories that included channel unit, complexity, cover, disturbance, riparian, size, substrate, temperature, water quality, and woody debris. Within each category, metrics were ranked according to their MIC value (Figure 1). Our strategy was to select one or two variables with the highest MIC score within each category so that covariates describe different aspects of rearing habitat (e.g., substrate, temperature, etc.). 

```{r MINE_stats}

# # what are some possible habitat covariates?
# poss_hab_mets = hab_dict %>%
#   filter(MetricCategory != 'Categorical') %>%
#   filter(ShortName %in% names(fish_hab)) %>%
#   pull(ShortName)

# pull from which CHaMP metrics could be generated from DASH
poss_hab_mets = read_csv('data/prepped/hab_dict_champ_dash.csv') %>%
  select(-starts_with('X')) %>%
  filter(DASH_gen) %>%
  filter(MetricCategory != 'Categorical') %>%
  pull(ShortName) %>%
  unique()
poss_hab_mets = c(poss_hab_mets, 'aug_avg_temp', 'avg_aug_temp')

mine_res = fish_hab %>%
  split(list(.$Species)) %>%
  map_df(.id = 'Species',
         .f = function(x) {
           x %>%
             mutate(fish_dens = log(fish_dens + 0.005)) %>%
             estimate_MIC(covars = poss_hab_mets,
                          response = 'fish_dens')
         }) %>%
  left_join(hab_dict %>%
              filter(MetricGroupName == 'Visit Metric') %>%
              select(Metric = ShortName,
                     MetricCategory,
                     Name),
            by = 'Metric') %>%
  # put the metric names in descending order by MIC
  mutate_at(vars(Metric, Name),
            list(~ fct_reorder(., .x = MIC))) %>%
  select(Species, MetricCategory, Metric, everything()) %>%
  arrange(Species, MetricCategory, desc(MIC))

mine_plot_df = mine_res #%>%
  # filter out some metrics with too many NAs or 0s
  # filter((perc_NA < 0.2 & non_0 > 100) | MetricCategory == 'Temperature') #%>%
  # # filter out metrics with very low variance
  # filter(var < 0.1) %>% select(1:11)
  # # filter out area and volume metrics
  # filter(!grepl('Area', Metric),
  #        !grepl('Vol', Metric),
  #        Metric != 'Lgth_Wet')
```

```{r hab_mets_correlation}

library(corrr)
#----------------------------------------------
# Look at correlations between habitat metrics
#----------------------------------------------
# top metrics
sel_mets = poss_hab_mets
sel_mets = mine_plot_df %>%
  group_by(MetricCategory) %>%
  slice(1:5) %>%
  ungroup() %>%
  pull(Metric) %>%
  unique() %>%
  as.character()


corr_mat = hab_avg %>%
  select(one_of(sel_mets)) %>%
  corrr::correlate()

corr_mat %>%
  rearrange(absolute = F) %>%
  shave(upper = T) %>% 
  stretch() %>%
  filter(!is.na(r)) %>%
  filter(abs(r) > 0.5)

corr_mat %>%
  # rearrange(absolute = F) %>%
  shave(upper = T) %>% 
  rplot(legend = T,
        print_cor = T)

network_plot(corr_mat)
  
```


## Results

```{r mine_plot}
# make a plot of MIC values for all species
mine_p = mine_plot_df %>%
  ggplot(aes(x = Name,
             y = MIC,
             fill = Species)) +
  geom_col(position = position_dodge(1)) +
  coord_flip() +
  facet_wrap(~ MetricCategory,
             scales = 'free_y',
             ncol = 3) +
  scale_fill_brewer(palette = 'Set1',
                    guide = guide_legend(nrow = 1)) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 5))

mine_p

```

```{r}
mine_p2 = mine_plot_df %>%
  ggplot(aes(x = Name,
             y = MIC,
             fill = Species)) +
  geom_col(position = position_dodge(1)) +
  coord_flip() +
  scale_fill_brewer(palette = 'Set1',
                    guide = guide_legend(nrow = 1)) +
  theme(legend.position = 'bottom',
        axis.text = element_text(size = 5))

mine_p2
```

In the end, we decided to use the same metrics for both species, and chose the following metrics:

```{r}
# sel_hab_mets = crossing(Species = c('Chinook', 
#                                     'Steelhead'),
#                         Metric = c('CU_Freq',
#                                    'DpthThlwg_UF_CV',
#                                    'WetWDRat_Avg',
#                                    'FishCovTotal',
#                                    'DistPrin1',
#                                    'RipCovGrnd',
#                                    'DpthThlwg_Avg',
#                                    'SubEstGrvl',
#                                    'SubD50',
#                                    'SubLT6',
#                                    'AvgHourly',
#                                    'Max7dAM',
#                                    'Cond',
#                                    'LWFreq_Wet'))
# 
# 
# sel_hab_mets = crossing(Species = c('Chinook', 
#                                     'Steelhead'),
#                         Metric = poss_hab_mets)

# based on conversation with Mike and Richie
sel_hab_mets = crossing(Species = c('Chinook', 
                                    'Steelhead'),
                        Metric = c('UcutArea_Pct',
                                   'FishCovNone',
                                   'SubEstGrvl',
                                   'FstTurb_Freq',
                                   'FstNT_Freq',
                                   'CU_Freq',
                                   'SlowWater_Pct',
                                   'NatPrin1',
                                   'DistPrin1',
                                   'avg_aug_temp',
                                   'Sin_CL',
                                   'WetWdth_CV',
                                   'WetBraid',
                                   'WetSC_Pct',
                                   # 'Discharge',
                                   'WetWdth_Int',
                                   'LWFreq_Wet',
                                   'LWVol_WetFstTurb'))


```


# Fit QRF Models

## Methods

Random forest models cannot handle missing data by default, so the first step is to impute any missing data. In order to limit the amount of imputation necessary, we first delete any observation with too many missing values, and then impute the rest. 

```{r impute_qrf_data}
qrf_mod_df = fish_hab %>%
  split(list(.$Species)) %>%
  map_df(.id = 'Species',
         .f = function(x) {
    spp = unique(x$Species)
    
    covars = sel_hab_mets %>%
      filter(Species == spp) %>%
      pull(Metric)
    
    data = impute_missing_data(data = x %>%
                                 mutate_at(vars(Watershed, Year),
                                           list(fct_drop)),
                               covars = covars,
                               impute_vars = c('Watershed', 'Elev_M', 'Sin', 'Year', 'CUMDRAINAG'),
                               method = 'missForest') %>%
      select(Site, Watershed, Year, LON_DD, LAT_DD, fish_dens, VisitID, one_of(covars))
    
    return(data)

  })

```

When fitting the QRF model, we would like to accomodate actual 0's so we offset the density by a small positive value. We then log-transformed the resulting fish density, and used that as the response.

```{r fit_qrf_model}
# set the density offset (to accommodate 0z)
dens_offset = 0.005

# fit random forest models
set.seed(4)
qrf_mods = qrf_mod_df %>%
  split(list(.$Species)) %>%
  map(.f = function(z) {
    
    covars = sel_hab_mets %>%
      filter(Species == unique(z$Species)) %>%
      pull(Metric)
    
    set.seed(3)
    qrf_mod = quantregForest(x = z %>%
                               select(one_of(covars)) %>%
                               as.matrix,
                             y = z %>%
                               mutate_at(vars(fish_dens),
                                         list(~ log(. + dens_offset))) %>%
                               select(fish_dens) %>%
                               as.matrix(),
                             keep.inbag = T,
                             ntree = 1000)
    
    return(qrf_mod)
  })

```

```{r relative_importance plot}

rel_imp_p = qrf_mods %>%
  map(.f = function(x) {
    as_tibble(x$importance,
              rownames = 'Metric') %>%
      mutate(relImp = IncNodePurity / max(IncNodePurity)) %>%
      left_join(hab_dict %>%
                  select(Metric = ShortName,
                         Name)) %>%
      mutate_at(vars(Metric, Name),
                list(~ fct_reorder(., relImp))) %>%
      arrange(Metric) %>%
      distinct() %>%
      ggplot(aes(x = Name,
                 y = relImp)) +
      geom_col(fill = 'gray40') +
      coord_flip() +
      labs(x = 'Metric',
           y = 'Relative Importance')
      
  })

ggpubr::ggarrange(plotlist = rel_imp_p,
                  nrow = 1,
                  ncol = 2,
                  labels = names(qrf_mods))

```

```{r partial_dependence_plots}
# for Chinook
chnk_pdp = plot_partial_dependence(qrf_mods[[1]],
                                   qrf_mod_df %>%
                                     filter(Species == 'Chinook'),
                                   hab_dict,
                                   scales = 'free_x')

# for steelhead
sthd_pdp = plot_partial_dependence(qrf_mods[[2]],
                                   qrf_mod_df %>%
                                     filter(Species == 'Steelhead'),
                                   hab_dict,
                                   scales = 'free_x')


```

```{r}
chnk_pdp
```

```{r}
sthd_pdp
```


# Predict at Habitat Sites

```{r}
# what quantile is a proxy for capacity?
pred_quant = 0.9

wtsd = "Lemhi"
wtsd = 'Upper Grande Ronde'
wtsd = c('Upper Grande Ronde', 'Minam')
# wtsd = 'John Day'
qrf_mod_df %>%
  filter(Species == 'Chinook',
         Watershed %in% wtsd) %>%
  left_join(champ_dash %>%
              filter(Watershed %in% wtsd) %>%
              group_by(Site, StreamName) %>%
              summarise_at(vars(Lgth_Wet, Area_Wet),
                           list(mean))) %>%
  mutate(pred_cap = predict(qrf_mods[[1]],
                            newdata = select(., one_of(unique(sel_hab_mets$Metric))),
                            what = pred_quant)) %>%
  mutate(pred_cap = exp(pred_cap) - dens_offset,
         pred_cap = pred_cap * Lgth_Wet / Area_Wet) %>%
  st_as_sf(coords = c('LON_DD', 'LAT_DD'),
           crs = 4326) %>%
  st_transform(crs = 5070) %>%
  ggplot() +
  geom_sf(aes(color = pred_cap)) +
  scale_color_viridis_c(direction = -1) +
  theme(axis.text = element_blank()) +
  labs(color = expression(Parr / m^2),
       title = paste(wtsd, collapse = ' '))

```



## Methods

## Results

# Extrapolation to Other Sites

To estimate watershed capacities, we used QRF predictions at all the sites where we had habitat data, and used those as the response variable in an extrapolation model. The covariates for this model are globally available attributes (GAAs) which are attached to every master sample point.

```{r}
covars = unique(sel_hab_mets$Metric)
hab_impute = hab_avg %>%
  mutate_at(vars(Watershed, Channel_Type),
            list(fct_drop)) %>%
  impute_missing_data(data = .,
                      covars = covars,
                      impute_vars = c('Watershed', 
                                      'Elev_M', 
                                      'Channel_Type', 
                                      'CUMDRAINAG'),
                      method = 'missForest') %>%
  select(Site, Watershed, LON_DD, LAT_DD, VisitYear, Lgth_Wet, Area_Wet, one_of(covars))

pred_hab_sites = hab_impute %>%
  mutate(chnk_per_m = predict(qrf_mods[[1]],
                              newdata = select(., one_of(unique(sel_hab_mets$Metric))),
                              what = pred_quant),
         chnk_per_m = exp(chnk_per_m) - dens_offset,
         chnk_per_m2 = chnk_per_m * Lgth_Wet / Area_Wet)

# filter out sites outside the Chinook domain
data("chnk_domain")
snap_dist = 1000

pred_hab_sites_chnk = pred_hab_sites %>%
  filter(!is.na(LON_DD)) %>%
  st_as_sf(coords = c('LON_DD', 'LAT_DD'),
           crs = 4326) %>%
  st_transform(crs = st_crs(chnk_domain)) %>%
  as_Spatial() %>%
  maptools::snapPointsToLines(chnk_domain %>%
                                mutate(id = 1:n()) %>%
                                select(id, MPG) %>%
                                as_Spatial(),
                              maxDist = snap_dist,
                              withAttrs = T,
                              idField = 'id') %>%
  as('sf') %>%
  as_tibble() %>%
  select(-nearest_line_id, -snap_dist, -geometry) %>%
  # add sites in the John Day
  bind_rows(st_read('data/raw/domain/Chnk_JohnDay_TrueObs.shp') %>%
              as_tibble() %>%
              select(Site) %>%
              inner_join(pred_hab_sites))

```

```{r}
#----------------------------------------
# pull in survey design related data
#----------------------------------------
# Calculate GRTS design weights.

# pull in info about what strata each CHaMP site was assigned to (using 2014 as reference year)
site_strata = pred_hab_sites_chnk %>%
  select(Site, Watershed) %>%
  distinct() %>%
  left_join(gaa %>%
              select(Site,
                     # CHaMPsheds,
                     strata = AStrat2014)) %>%
  mutate(site_num = str_split(Site, '-', simplify = T)[,2]) %>%
  mutate(strata = if_else(Watershed == 'Asotin',
                          site_num,
                          if_else(Watershed == 'Entiat' & grepl('ENT00001', Site),
                                  paste('EntiatIMW', site_num, sep = '_'),
                                  strata))) %>%
  mutate(strata = if_else(grepl('EntiatIMW', strata),
                          str_remove(strata, '[[:digit:]]$'),
                          strata),
         strata = if_else(grepl('EntiatIMW', strata),
                          str_remove(strata, '[[:digit:]]$'),
                          strata)) %>%
  filter(!is.na(strata)) %>%
  mutate(strata = paste(Watershed, strata, sep = '_')) %>%
  select(-site_num)

# read in data from the CHaMP frame
champ_frame_df = read_csv('data/prepped/champ_frame_data.csv') %>%
  mutate(Target2014 = ifelse(is.na(AStrat2014), 'Non-Target', Target2014)) %>%
  mutate(AStrat2014 = ifelse(AStrat2014 == 'Entiat IMW', paste('EntiatIMW', GeoRchIMW, sep = '_'), AStrat2014)) %>%
  filter(Target2014 == 'Target') %>%
  rename(Watershed = CHaMPshed)
  
# what strata do we have?
frame_strata = champ_frame_df %>%
  mutate(strata = paste(Watershed, AStrat2014, sep='_')) %>%
  select(Watershed, 
         strata) %>%
  distinct()

# how long is each strata?
chnk_strata_length = champ_frame_df %>%
  filter(!is.na(UseTypCHSP)) %>%
  mutate(strata = paste(Watershed, AStrat2014, sep='_')) %>%
  select(Watershed, matches("Strat"), FrameLeng) %>%
  group_by(Watershed, strata) %>%
  summarise(tot_length_km = sum(FrameLeng) / 1000) %>%
  ungroup() %>%
  mutate_at(vars(Watershed, strata), 
            list(as.factor)) %>%
  arrange(Watershed, strata)

sthd_strata_length = champ_frame_df %>%
  filter(!is.na(UseTypSTSU)) %>%
  mutate(strata = paste(Watershed, AStrat2014, sep='_')) %>%
  select(Watershed, matches("Strat"), FrameLeng) %>%
  group_by(Watershed, strata) %>%
  summarise(tot_length_km = sum(FrameLeng) / 1000) %>%
  bind_rows(tibble(Watershed = 'Asotin',
                   strata = paste('Asotin', c('CC', 'NF', 'SF'), sep = '_'),
                   tot_length_km = 12)) %>%
  ungroup() %>%
  mutate_at(vars(Watershed, strata), 
            list(as.factor)) %>%
  arrange(Watershed, strata)


# how many sites in each strata? and what is the length of each strata?
strata_tab = pred_hab_sites_chnk %>%
  select(Site, Watershed, matches('per_m')) %>%
  left_join(site_strata) %>%
  filter(strata != 'Entiat_Entiat IMW') %>%
  mutate_at(vars(Watershed),
            list(fct_drop)) %>%
  group_by(Watershed, strata) %>%
  summarise(n_sites = n_distinct(Site)) %>%
  ungroup() %>%
  full_join(chnk_strata_length) %>%
  mutate(n_sites = if_else(is.na(n_sites),
                           as.integer(0),
                           n_sites)) %>%
  # calculate the weight of each site in each strata
  mutate(site_weight = if_else(n_sites > 0,
                               tot_length_km / n_sites,
                               as.numeric(NA)))


strata_test = frame_strata %>%
  full_join(strata_tab) %>%
  mutate_at(vars(Watershed),
            list(fct_drop)) %>%
  mutate(n_sites = if_else(is.na(n_sites),
                           as.integer(0),
                           n_sites))

# what frame strata don't have any sites in them?
strata_test %>%
  filter(n_sites == 0,
         !is.na(tot_length_km)) %>%
  arrange(Watershed, strata) %>%
  as.data.frame()


# what strata that we have sites for are not in the frame strata?
strata_test %>%
  filter(n_sites > 0,
         (is.na(tot_length_km) |
            tot_length_km == 0)) %>%
  as.data.frame()


strata_test %>%
  group_by(Watershed) %>%
  summarise_at(vars(tot_length_km),
               list(sum),
               na.rm = T) %>%
  left_join(strata_test %>%
              filter(n_sites == 0) %>%
              group_by(Watershed) %>%
              summarise_at(vars(missing_length = tot_length_km),
                           list(sum),
                           na.rm = T)) %>%
  mutate_at(vars(missing_length),
            list(~ if_else(is.na(.), 0, .))) %>%
  mutate(perc_missing = missing_length / tot_length_km) %>%
  mutate_at(vars(perc_missing),
            list(~ if_else(is.na(.), 0, .))) %>%
  arrange(desc(perc_missing))



```


```{r}
data(gaa)

gaa_all = gaa %>%
  rename(Channel_Type = ChanlType)

# possible covariates from among the GAAs
gaa_covars = c('TRange', 'GDD', 'Precip', 'Elev_M', 'CHaMPsheds', 'NatPrin1', 'NatPrin2', 'DistPrin1', 'BFW_M', 'SrtCumDrn', 'StrmPwr', 'Slp_NHD_v1', 'Channel_Type', 'MAVELV', 'WIDE_BF')

# what type of covariate is each GAA?
gaa_class = gaa_all %>%
  select(one_of(gaa_covars)) %>%
  as.list() %>%
  map_chr(.f = class)

# which ones are numeric?
gaa_num = names(gaa_class)[gaa_class %in% c('integer', 'numeric')]
# which ones are categorical?
gaa_catg = names(gaa_class)[gaa_class %in% c('factor', 'character')]

# compare range of covariates from model dataset and prediction dataset
range_comp = bind_rows(gaa_all %>%
                         filter(!Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'non-DASH Sites'),
                       gaa_all %>%
                         filter(Site %in% unique(pred_hab_sites$Site)) %>%
                         select(Site,
                                one_of(gaa_num)) %>%
                         distinct() %>%
                         gather(Metric, value, -Site) %>%
                         mutate(Source = 'DASH Sites')) %>%
  mutate_at(vars(Source, Metric),
            list(as.factor))

range_max = range_comp %>%
  group_by(Metric, Source) %>%
  summarise_at(vars(value),
               tibble::lst(min, max),
               na.rm = T) %>%
  filter(Source == 'DASH Sites') %>%
  ungroup() %>%
  gather(type, value, -Metric, -Source)

covar_range_p = range_comp %>%
  ggplot(aes(x = Source,
             y = value,
             fill = Source)) +
  geom_boxplot() +
  facet_wrap(~ Metric,
             scales = 'free') +
  geom_hline(data = range_comp %>%
               group_by(Metric, Source) %>%
               summarise_at(vars(value),
                            tibble::lst(min, max),
                            na.rm = T) %>%
               filter(Source == 'DASH Sites') %>%
               ungroup() %>%
               gather(type, value, min, max),
             aes(yintercept = value),
             lty = 2,
             color = 'darkgray') +
  theme_minimal()

covar_range_p

# look at correlations among GAAs
library(corrr)
corr_df = gaa_all %>%
  select(one_of(gaa_num)) %>%
  correlate()

corr_df %>%
  shave(upper = T) %>%
  stretch() %>%
  arrange(desc(abs(r))) %>%
  filter(abs(r) >= 0.5)

corr_df %>%
  shave(upper = T) %>%
  rplot(legend = T,
        colors = c('blue', 'white', 'indianred2'),
        print_cor = T)

corr_mat = corr_df %>%
  select(-1) %>%
  as.matrix() 
rownames(corr_mat) = corr_df$rowname

ggcorrplot(corr_mat,
           lab = T,
           lab_size = 2,
           hc.order = T,
           colors = c('#2166AC', 'white', '#B2182B'),
           tl.cex = 8,
           type = 'full')
```


```{r}
# reduce number of covariates bases on correlations
gaa_covars = c('TRange', 
               # 'GDD', 
               # 'Precip',
               'Elev_M', 
               'CHaMPsheds', 
               'NatPrin1', 
               # 'NatPrin2', 
               'DistPrin1', 
               # 'BFW_M', 
               'SrtCumDrn', 
               'StrmPwr',
               'Slp_NHD_v1', 
               'Channel_Type', 
               # 'MAVELV', 
               'WIDE_BF')

# what type of covariate is each GAA?
gaa_class = gaa_all %>%
  select(one_of(gaa_covars)) %>%
  as.list() %>%
  map_chr(.f = class)

# which ones are numeric?
gaa_num = names(gaa_class)[gaa_class %in% c('integer', 'numeric')]
# which ones are categorical?
gaa_catg = names(gaa_class)[gaa_class %in% c('factor', 'character')]

# Center the covariates
# filter out sites with covariates outside range of covariates used to fit extrapolation model
out_range_sites = gaa_all %>%
  # filter out a few areas
  filter(!HUC6NmNRCS %in% c('Upper Sacramento', 'Southern Oregon Coastal', 'Puget Sound', 'Northern California Coastal', 'Oregon Closed Basins')) %>%
  # don't use AEM sites in model
  filter(!grepl('^AEM', Site)) %>%
  select(one_of(gaa_num), Site) %>%
  gather(Metric, value, -Site) %>%
  left_join(select(range_max, -Source) %>%
              spread(type, value)) %>%
  group_by(Metric) %>%
  filter(value > max |
           value < min) %>%
  ungroup() %>%
  pull(Site) %>%
  unique()

# center covariates
gaa_summ = inner_join(pred_hab_sites %>%
                          select(Site) %>%
                          distinct(),
                        gaa_all %>%
                          select(Site, one_of(gaa_num))) %>%
  gather(GAA, value, -Site) %>%
  group_by(GAA) %>%
  summarise(metric_mean = mean(value, na.rm=T),
            metric_sd = sd(value, na.rm=T)) %>%
  ungroup()

# extrapolation model data set, with normalized covariates
mod_data = inner_join(pred_hab_sites_chnk %>%
                        select(Site:avg_aug_temp,
                               matches('per_m')),
                      gaa_all %>%
                        select(Site, one_of(gaa_num))) %>%
  gather(GAA, value, one_of(gaa_num)) %>%
  left_join(gaa_summ) %>%
  mutate(norm_value = (value - metric_mean) / metric_sd) %>%
  select(-(value:metric_sd)) %>%
  spread(GAA, norm_value) %>%
  left_join(gaa_all %>%
              select(Site, one_of(gaa_catg)))

mod_data %<>%
  bind_cols(mod_data %>%
              is.na() %>%
              as_tibble() %>%
              select(one_of(gaa_covars)) %>%
              transmute(n_NA = rowSums(.))) %>%
  filter(n_NA == 0) %>%
  mutate_at(vars(Watershed, CHaMPsheds, Channel_Type),
            list(fct_drop)) 

```

```{r}
# calculate adjusted weights for all predicted QRF capacity sites
mod_data_weights = mod_data %>%
  left_join(site_strata) %>%
  left_join(strata_tab) %>%
  filter(!is.na(site_weight)) %>%
  group_by(Watershed) %>%
  mutate(sum_weights = sum(site_weight)) %>%
  ungroup() %>%
  mutate(adj_weight = site_weight / sum_weights)

# split by response: per meter or per meter sq
mod_data_list = mod_data_weights %>%
  gather(response, qrf_cap, matches('per_m')) %>%
  select(-(n_sites:sum_weights)) %>%
  split(list(.$response))


#-------------------------------------------------------------
# Set up the survey design.

# getOption('survey.lonely.psu')
# this will prevent strata with only 1 site from contributing to the variance
# options(survey.lonely.psu = 'certainty')
# this centers strata with only 1 site to the sample grand mean; this is conservative
options(survey.lonely.psu = 'adjust')

surv_design = mod_data_list %>%
  map(.f = function(x) {
    svydesign(id = ~ 1,
              data = x,
              # strata = ~ Watershed,
              strata = ~ strata,
              weights = ~ adj_weight)
  })

```

```{r}
full_form = as.formula(paste('log(qrf_cap) ~ -1 + (', paste(gaa_covars, collapse = ' + '), ')'))

model_svy_df = mod_data_weights %>%
  gather(response, qrf_cap, matches('per_m')) %>%
  select(-(n_sites:sum_weights)) %>%
  group_by(response) %>%
  nest() %>%
  mutate(design = map(data,
                      .f = function(x) {
                        svydesign(id = ~ 1,
                                  data = x,
                                  strata = ~ Watershed,
                                  # strata = ~ strata,
                                  weights = ~ adj_weight)
                      })) %>%
  mutate(mod_full = map(design,
                        .f = function(x) {
                          svyglm(full_form,
                                 design = x)
                        }),
         mod_no_champ = map(design,
                            .f = function(x) {
                              svyglm(update(full_form, .~ . -CHaMPsheds),
                                     design = x)
                            }))

```



```{r}

full_form = as.formula(paste('log(qrf_cap) ~ -1 + (', paste(gaa_covars, collapse = ' + '), ')'))
# extrap_mods = mod_data_list %>%
#   map(.f = function(x) {
#     mod_full = lm(full_form,
#                   data = x,
#                   na.action = na.fail)
#     
#     dd = dredge(mod_full)
#     mod_best = get.models(dd,
#                           subset = delta == 0)[[1]]
#     mod_avg = model.avg(dd,
#                         subset = cumsum(weight) < 0.96)
#   })

model_df = mod_data %>%
  filter(n_NA == 0) %>%
  gather(response, qrf_cap, matches('per_m')) %>%
  group_by(response) %>%
  nest() %>%
  mutate(mod_full = map(data,
                        .f = function(x) {
                          lm(full_form,
                             data = x)
                        }),
         mod_no_champ = map(data,
                        .f = function(x) {
                          lm(update(full_form, .~ . -CHaMPsheds),
                             data = x)
                        }))



# where do we want to make extrapolation predictions?
gaa_pred = gaa_all %>%
  # filter out a few areas
  filter(!HUC6NmNRCS %in% c('Upper Sacramento', 'Southern Oregon Coastal', 'Puget Sound', 'Northern California Coastal', 'Oregon Closed Basins')) %>%
  # don't use AEM sites in model
  filter(!grepl('^AEM', Site)) %>%
  # don't use non-GRTS sites
  filter(SiteID_alt != 'NonGRTSSite' | is.na(SiteID_alt)) %>%
  filter(!grepl('mega', Site, ignore.case = T)) %>%
  # note which sites have GAAs outside range of CHaMP sites GAAs
  mutate(inCovarRange = ifelse(Site %in% out_range_sites, F, T)) %>%
  select(Site, one_of(gaa_covars), Lon, Lat, inCovarRange, HUC6NmNRCS, HUC8NmNRCS, HUC10NmNRC, HUC12NmNRC, chnk) %>%
  gather(GAA, value, one_of(gaa_num)) %>%
  left_join(gaa_summ) %>%
  mutate(norm_value = (value - metric_mean) / metric_sd) %>%
  select(-(value:metric_sd)) %>%
  spread(GAA, norm_value)
  
y = gaa_pred %>%
  filter(CHaMPsheds %in% unique(model_df$data[[1]]$Watershed)) %>%
  select(Site, CHaMPsheds, one_of(gaa_covars)) %>%
  na.omit() %>%
  left_join(gaa_pred) %>%
  mutate_at(vars(Channel_Type),
            list(as.factor)) %>%
  bind_cols(predict(model_df$mod_full[[1]],
                    newdata = .,
                    se.fit = F,
                    interval = 'confidence',
                    type = 'response') %>%
              as_tibble() %>%
              select(chnk_per_m = fit,
                     chnk_per_m_lwr = lwr,
                     chnk_per_m_upr = upr)) %>%
  bind_cols(predict(model_df$mod_full[[2]],
                    newdata = .,
                    se.fit = F,
                    interval = 'confidence',
                    type = 'response') %>%
              as_tibble() %>%
              select(chnk_per_m2 = fit,
                     chnk_per_m2_lwr = lwr,
                     chnk_per_m2_upr = upr)) %>%
  mutate_at(vars(matches('per_m')),
            list(exp))

```

```{r}
wtsd = 'Lemhi'
wtsd = 'South Fork Salmon'
wtsd = 'Lower Snake-Tucannon'
wtsd = 'Wenatchee'
wtsd = 'Methow'

y %>%
  filter(HUC8NmNRCS %in% wtsd) %>%
  filter(chnk == 'Yes') %>%
  st_as_sf(coords = c('Lon', 'Lat'),
           crs = 4326) %>%
  st_transform(crs = 5070) %>%
  ggplot() +
  geom_sf(aes(color = chnk_per_m2)) +
  scale_color_viridis_c(direction = -1) +
  theme(axis.text = element_blank()) +
  labs(color = expression(Parr / m^2),
       title = paste(wtsd, collapse = ' '))
  
```


# References