---
title: "Re-Creating IRA Capacity Deficits"
author:
  - name: Kevin See
    affiliation: biomark
    email: kevin.see@merck.com
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  bookdown::html_document2:
    theme: simplex
    toc: yes
    toc_depth: 3
    toc_float: yes
    fig_height: 8
    fig_width: 8
    number_sections: no
    pandoc_args:
    - --lua-filter=templates/scholarly-metadata.lua
    - --lua-filter=templates/author-info-blocks.lua
    - --lua-filter=templates/pagebreak.lua
  bookdown::pdf_document2:
    pandoc_args:
    - --lua-filter=templates/scholarly-metadata.lua
    - --lua-filter=templates/author-info-blocks2.lua
    - --lua-filter=templates/pagebreak.lua
    fig_height: 7.5
    fig_width: 6
    number_sections: no
    includes:
      in_header: "templates/header_ABS.tex"
  bookdown::word_document2: 
    fig_caption: yes
    fig_height: 7
    fig_width: 6
    toc: yes
    number_sections: no
    pandoc_args:
    - --lua-filter=templates/scholarly-metadata.lua
    - --lua-filter=templates/author-info-blocks.lua
    - --lua-filter=templates/pagebreak.lua
    reference_docx: "templates/ReportTemplate.docx"
institute:
- biomark: Biomark, Inc.
csl: "templates/american-fisheries-society.csl"
# bibliography: references.bib
bibliography: 
  - references.bib
  - packages.bib
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# setwd('reports')
# load knitr for markdown
library(knitr)
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE,
                      # error = FALSE,
                      message=FALSE)
#options(tinytex.verbose = TRUE)
options(knitr.kable.NA = '-')
# options(knitr.table.format = "pandoc")

library(kableExtra)
```

```{r}
# load needed packages
library(QRFcapacity)
library(tidyverse)
library(magrittr)
library(sf)
library(janitor)
library(ggpubr)
library(readxl)
# library(ggrepel)
# library(scales)
# library(ggspatial)

# theme_set(theme_bw())
theme_set(theme_pubr(base_size = 10))
```

```{r package-bibtex, eval = F}
knitr::write_bib(c("base",
                   "survey", 
                   "sf", 
                   "knitr", 
                   "rmarkdown"),
                 file = 'packages.bib')
```

```{r ira-boundaries}
# Chinook
data("rch_200")
ira_chnk_list = list("East Fork Salmon" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - East Fork Salmon River",
                         chnk),
                "Valley Creek" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Valley Creek",
                         chnk),
                "Yankee Fork" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Yankee Fork",
                         chnk),
                "North Fork Salmon" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - North Fork Salmon River",
                         GNIS_Name %in% c("Dahlonega Creek",
                                          "Hughes Creek",
                                          # "Moose Creek",
                                          "North Fork Salmon River",
                                          "Sheep Creek",
                                          "Twin Creek"),
                         chnk),
                "Lemhi" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Lemhi River",
                         chnk,
                         HUC8_code == "17060204"),
                "Pahsimeroi" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Pahsimeroi River",
                         chnk),
                "Panther Creek" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Panther Creek",
                         GNIS_Name %in% c("Big Deer Creek",
                                          "Blackbird Creek",
                                          "Clear Creek",
                                          "Moyer Creek",
                                          "Musgrove Creek",
                                          "Napias Creek",
                                          "Panther Creek"),
                         chnk),
                "Upper Salmon" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Salmon River Upper Mainstem above Redfish Lake",
                         chnk))

wtsd_chnk_poly = ira_chnk_list %>%
  map(.f = function(x) {
    x %>%
      st_union() %>%
      st_convex_hull()
  })

# get Chinook range from 200m reaches
chnk_wtsd_range = ira_chnk_list %>%
  map(.f = function(x) {
    x %>%
      mutate(Species = "Chinook Salmon",
             SciName = "Oncorhynchus tshawytscha") %>%
      select(StreamName = GNIS_Name,
             Species, SciName,
             UseType = chnk_use,
             ESU_DPS = chnk_ESU_DPS,
             MPG = chnk_MPG,
             NWR_POPID = chnk_NWR_POPID,
             NWR_NAME = chnk_NWR_NAME)
  })
chnk_domain = do.call(rbind, chnk_wtsd_range)

# Steelhead
ira_sthd_list = list("East Fork Salmon" = rch_200 %>%
                  filter(sthd_NWR_NAME == "Steelhead (Snake River Basin DPS) - East Fork Salmon River",
                         sthd),
                "Valley Creek" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Valley Creek",
                         sthd),
                "Yankee Fork" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Yankee Fork",
                         sthd),
                "North Fork Salmon" = rch_200 %>%
                  filter(sthd_NWR_NAME == "Steelhead (Snake River Basin DPS) - North Fork Salmon River",
                         GNIS_Name %in% c("Dahlonega Creek",
                                          "Hughes Creek",
                                          # "Moose Creek",
                                          "North Fork Salmon River",
                                          "Sheep Creek",
                                          "Twin Creek"),
                         sthd),
                "Lemhi" = rch_200 %>%
                  filter(sthd_NWR_NAME == "Steelhead (Snake River Basin DPS) - Lemhi River",
                         sthd,
                         HUC8_code == "17060204"),
                "Pahsimeroi" = rch_200 %>%
                  filter(sthd_NWR_NAME == "Steelhead (Snake River Basin DPS) - Pahsimeroi River",
                         GNIS_Name %in% c("Sulphur Creek",
                                           "Pahsimeroi River", 
                                           "Patterson Creek"),
                         sthd),
                "Panther Creek" = rch_200 %>%
                  filter(sthd_NWR_NAME == "Steelhead (Snake River Basin DPS) - Panther Creek",
                         GNIS_Name %in% c("Big Deer Creek",
                                          "Blackbird Creek",
                                          "Clear Creek",
                                          "Moyer Creek",
                                          "Musgrove Creek",
                                          "Napias Creek",
                                          "Panther Creek"),
                         sthd),
                "Upper Salmon" = rch_200 %>%
                  filter(chnk_NWR_NAME == "Chinook Salmon (Snake River Spring/Summer-run ESU) - Salmon River Upper Mainstem above Redfish Lake",
                         sthd))

wtsd_sthd_poly = ira_sthd_list %>%
  map(.f = function(x) {
    x %>%
      st_union() %>%
      st_convex_hull()
  })

# get steelhead range from 200m reaches
sthd_wtsd_range = ira_sthd_list %>%
  map(.f = function(x) {
    x %>%
      mutate(Species = "steelhead",
             SciName = "Oncorhynchus mykiss") %>%
      select(StreamName = GNIS_Name,
             Species, SciName,
             UseType = sthd_use,
             ESU_DPS = sthd_ESU_DPS,
             MPG = sthd_MPG,
             NWR_POPID = sthd_NWR_POPID,
             NWR_NAME = sthd_NWR_NAME)
  })
sthd_domain = do.call(rbind, sthd_wtsd_range)

```


```{r upper-lemhi-bndry}
library(nhdplusTools)

hayden_comid = tibble(site = "Hayden Confluence",
                      lat = 44.869942,
                      lon = -113.626182) %>%
    st_as_sf(coords = c("lon", "lat"),
             crs = 4326) %>%
    nhdplusTools::discover_nhdplus_id()

upper_lem = nhdplusTools::plot_nhdplus(outlets = list(hayden_comid),
                           streamorder = 2,
                           actually_plot = F)
upper_lem$basin = upper_lem$basin %>%
  nngeo::st_remove_holes() %>%
  st_transform(st_crs(chnk_domain))

```

```{r get-cap-all-models}
cap_df = c('juv_summer',
           # 'juv_summer_dash',
           'redds',
           "juv_winter") %>%
  as.list() %>%
  rlang::set_names() %>%
  map_df(.id = "lifestage",
         .f = function(x) {
           rch_cap = st_read(paste0("../output/gpkg/Rch_Cap_RF_", x,
                                    ".gpkg"),
                             quiet = T)
           
           ira_chnk_rch = ira_chnk_list %>%
             map_df(.id = "watershed",
                    .f = function(x) {
                      tibble(chnk_rch = list(x %>%
                                          inner_join(rch_cap %>%
                                                       select(-c(GNIS_Name:Watershed)) %>%
                                                       st_drop_geometry() %>%
                                                       as_tibble(),
                                                     by = "UniqueID")))
                    })
           
           ira_sthd_rch = ira_sthd_list %>%
             map_df(.id = "watershed",
                    .f = function(x) {
                      tibble(sthd_rch = list(x %>%
                                          inner_join(rch_cap %>%
                                                       select(-c(GNIS_Name:Watershed)) %>%
                                                       st_drop_geometry() %>%
                                                       as_tibble(),
                                                     by = "UniqueID")))
                    })
           
           
           if(x == "juv_summer") {
             rch_pts = st_read("../output/gpkg/Sum_Juv_Capacity.gpkg",
                               quiet = T) %>%
               st_transform(st_crs(rch_cap))
           } else if(x == "juv_summer_dash") {
             rch_pts = st_read("../output/gpkg/Sum_Juv_Capacity_DASH.gpkg",
                               quiet = T) %>%
               st_transform(st_crs(rch_cap))
           } else if(x == "redds") {
             rch_pts = st_read("../output/gpkg/Redds_Capacity.gpkg",
                               quiet = T) %>%
               st_transform(st_crs(rch_cap))
           } else if(x == "juv_winter") {
             rch_pts = st_read("../output/gpkg/Win_Juv_Capacity.gpkg",
                               quiet = T) %>%
               st_transform(st_crs(rch_cap))
           }
           
           ira_chnk_pts = wtsd_chnk_poly %>%
             map_df(.id = "watershed",
                    .f = function(x) {
                      tibble(chnk_pts = list(st_intersection(rch_pts, x)))
                    })
           
           ira_sthd_pts = wtsd_sthd_poly %>%
             map_df(.id = "watershed",
                    .f = function(x) {
                      tibble(sthd_pts = list(st_intersection(rch_pts, x)))
                    })
             
           ira_chnk_rch %>%
             full_join(ira_sthd_rch,
                       by = "watershed") %>%
             full_join(ira_chnk_pts,
                       by = "watershed") %>%
             full_join(ira_sthd_pts,
                       by = "watershed") %>%
             return()
         }) %>%
  left_join(wtsd_chnk_poly %>% 
              map_df(.id = "watershed",
                     .f = function(x) {
                       tibble(wtsd_chnk_poly = list(x))
                     }),
            by = "watershed") %>%
  left_join(wtsd_sthd_poly %>% 
              map_df(.id = "watershed",
                     .f = function(x) {
                       tibble(wtsd_sthd_poly = list(x))
                     }),
            by = "watershed")

cap_df %<>%
  pivot_longer(cols = -(lifestage:watershed),
               names_to = "name",
               values_to = "value") %>%
  mutate(Species = if_else(grepl('chnk', name), "Chinook",
                           if_else(grepl('sthd', name), "Steelhead",
                                   NA_character_))) %>%
  mutate(name = str_remove(name, "chnk_"),
         name = str_remove(name, "sthd_")) %>%
  pivot_wider(names_from = "name",
              values_from = "value") %>%
  left_join(tibble(Species = c("Chinook", "Steelhead"),
                   spp_domain = list(chnk_domain,
                                     sthd_domain)),
            by = "Species")

```

```{r current-capacity-est}
cap_est = cap_df %>%
  nest(inputs = rch:spp_domain) %>%
  mutate(cap_rch = map(inputs,
                       .f = function(y) {
                         calc_watershed_cap(wtsd_polygon = y$wtsd_poly[[1]],
                                            capacity_sf = y$rch[[1]],
                                            spp_range = y$spp_domain[[1]],
                                            by_stream = F) %>%
                           rename(n = n_rchs)
                       })) %>%
  mutate(cap_pts = map(inputs,
                       .f = function(y) {
                         calc_watershed_cap(wtsd_polygon = y$wtsd_poly[[1]],
                                            capacity_sf = y$pts[[1]],
                                            spp_range = y$spp_domain[[1]],
                                            by_stream = F) %>%
                           select(-area) %>%
                           rename(n = n_pts)
                       })) %>%
  pivot_longer(cols = c(cap_rch,
                        cap_pts),
               names_to = "source",
               values_to = "cap") %>%
  mutate(source = recode(source,
                         "cap_rch" = "Reaches",
                         "cap_pts" = "Points")) %>%
  unnest(cols = cap) %>%
  mutate(across(tot_length,
                ~ . / 1000))
  
cap_comp_df = cap_est %>%
  select(-inputs)
```

```{r recovery-goals}
params = read_excel('../../MRA_QRF/data/IRA/USIRA Tables (all Upper Salmon).xlsx',
                    "Escapement Data - Chinook",
                    range = "B2:C7") %>%
  add_column(Species = 'Chinook',
             .before = 0) %>%
  bind_rows(read_excel('../../MRA_QRF/data/IRA/USIRA Tables (all Upper Salmon).xlsx',
                    "Escapement Data - Steelhead",
                    range = "B2:C7") %>%
              add_column(Species = 'Steelhead',
                         .before = 0))


# set recovery goals
rec_goals = tibble(Watershed = c("Upper Salmon",
                                 "Valley Creek",
                                 "Yankee Fork",
                                 "East Fork Salmon",
                                 "Pahsimeroi",
                                 "Lemhi",
                                 "North Fork Salmon",
                                 "Panther Creek"),
                   Species = c("Chinook"),
                   Mean = c(656,
                            506,
                            248,
                            283,
                            402,
                            347,
                            NA,
                            NA),
                   Max = c(1419,
                           739,
                           343,
                           343,
                           822,
                           718,
                           NA,
                           NA),
                   MAT = c(1000,
                           500,
                           500,
                           1000,
                           1000,
                           2000,
                           500,
                           750)) %>%
  bind_rows(tibble(Watershed = c("Upper Salmon",
                                 "Valley Creek",
                                 "Yankee Fork",
                                 "East Fork Salmon",
                                 "Pahsimeroi",
                                 "Lemhi",
                                 "North Fork Salmon",
                                 "Panther Creek"),
                   Species = c("Steelhead"),
                   Mean = c(92,
                            193,
                            95,
                            30,
                            1156,
                            337,
                            252,
                            449),
                   Max = c(154,
                           278,
                           213,
                           54,
                           1614,
                           417,
                           349,
                           650),
                   MAT = c(483,
                           247,
                           270,
                           1000,
                           1000,
                           1000,
                           500,
                           500))) %>%
  mutate(`MAT+25%` = MAT * 1.25) %>%
  pivot_longer(cols = -c(Watershed:Species),
               names_to = "Scenario",
               values_to = "Escapement") %>%
  inner_join(params %>%
               mutate(Parameter = recode(Parameter,
                                         "Female Ratio" = "prop_fem",
                                         "Redds/Female" = "redd_per_fem",
                                         "Fecundity" = 'fecud',
                                         'Egg:Parr' = "egg_to_parr",
                                         "Parr:Presmolt" = "parr_to_presmolt")) %>%
               pivot_wider(id_cols = "Species",
                           names_from = "Parameter",
                           values_from = "Value"),
             by = "Species") %>%
  mutate(Redds = Escapement * prop_fem * redd_per_fem,
         Eggs = Redds * fecud,
         Parr = Eggs * egg_to_parr,
         Presmolts = Parr * parr_to_presmolt) %>%
  select(-c(prop_fem:parr_to_presmolt)) %>%
  pivot_longer(cols = c(Escapement, Redds:Presmolts), 
               names_to = "Lifestage",
               values_to = "Abundance")

```

```{r create-bar-plots-rch}

comp_df = cap_comp_df %>%
  rename(Lifestage = lifestage,
         Watershed = watershed) %>%
  mutate(Lifestage = recode(Lifestage,
                            "juv_summer" = "Parr",
                            "redds" = "Redds",
                            "juv_winter" = "Presmolts")) %>%
  mutate(Scenario = paste0("QRF_cap_", source)) %>%
  rename(Abundance = tot_cap) %>%
  select(any_of(names(rec_goals))) %>%
  bind_rows(rec_goals) %>%
  pivot_wider(names_from = "Scenario",
              values_from = "Abundance") %>%
  select(Species, Watershed, Lifestage, Mean:`MAT+25%`, starts_with("QRF")) %>%
  # mutate(cap_def = `MAT+25%` - QRF_cap,
  #        rel_def = cap_def / QRF_cap) %>%
  mutate(Lifestage = factor(Lifestage,
                            levels = c("Redds",
                                       "Parr",
                                       "Presmolts"))) %>%
  arrange(Watershed, Lifestage) %>%
  filter(!is.na(Lifestage)) %>%
  pivot_longer(cols = Mean:QRF_cap_Points,
               names_to = "scenario",
               values_to = "value") %>%
  mutate(scenario = factor(scenario,
                           levels = c("QRF_cap_Reaches",
                                      "QRF_cap_Points",
                                      "Mean",
                                      "Max",
                                      "MAT",
                                      "MAT+25%")),
         scenario = fct_rev(scenario))

bar_list_p = comp_df %>%
  split(list(.$Species, .$Lifestage)) %>%
  map(.f = function(df) {
    df %>%
      ggplot(aes(y = scenario,
                 x = value,
                 fill = scenario)) +
      geom_col(position = "dodge") +
      # scale_fill_brewer(palette = "Paired") +
      scale_fill_manual(values = c("QRF_cap_Reaches" = "black",
                                   "QRF_cap_Points" = "gray",
                                   "Mean" = "lightgreen",
                                   "Max" = "darkgreen",
                                   "MAT" = "lightblue",
                                   "MAT+25%" = "darkblue"),
                        name = "Abundance") +
      facet_wrap(~ Watershed,
                 scales = "free_x") +
      theme(legend.position = "bottom",
            axis.text.y = element_blank(),
            axis.title.y = element_blank(),
            axis.ticks.y = element_blank()) +
      labs(x = "Scenario",
           title = unique(df$Lifestage))
  })

```

# Introduction

When the IRA report came out, our QRF model included an extrapolation model based on GRTS master sample points. Since then, we have updated the extrapolation model to one that is based on a linear network of 200m reaches. This linear network is also based on an updated NHDPlus network, which often increases the length of various streams as the updated version has finer resolution. 

To examine the impact of this change in extrapolation models, we re-compiled capacity estimates at the watershed scale for all IRA watersheds, across 3 lifestages. Currently this has been done for Chinook. We computed capacity using the master sample points, and also using the 200m reaches. We used the updated NHDPlus network for both, so the total stream kilometers for each watershed is the same. For some small tributaries, there may not be a master sample point on that tributary within the range of that species. In such cases, we have no estimate of the carrying capacity of that tributary, so we set it to zero.

# Methods

We used the same life-history parameters used in the IRA addendum, shown in Table \@ref(tab:param-table).

```{r param-table}
params %>%
  kable(caption = "Life history parameters used to translate recovery goals into various life-stages.",
        digits = 2) %>%
  kable_styling()
```

# Results

The contemporary mean and max of recent escapements (translated to other lifestages), as well as the recovery goal of mean abundance threshold (MAT) and a 25% buffer on top of that (MAT+25%) are shown in the plots below, together with estimates of current capacity from master sample points and reaches. Estimates for both methods are shown in Table \@ref(tab:chnk-cap-tab) for Chinook, and Table \@ref(tab:sthd-cap-tab) for steelhead.

## Chinook

```{r chnk-parr-fig, fig.cap = "Various scenarios of Chinook parr abundance including QRF capacity estimates."}
bar_list_p$Chinook.Parr
```

```{r chnk-winter-fig, fig.cap = "Various scenarios of Chinook presmolt over-winter abundance including QRF capacity estimates."}
bar_list_p$Chinook.Presmolts
```

```{r chnk-redd-fig, fig.cap = "Various scenarios of Chinook redd abundance including QRF capacity estimates."}
bar_list_p$Chinook.Redds
```

```{r chnk-cap-tab}
cap_comp_df %>%
  filter(Species == "Chinook") %>%
  rename(Lifestage = lifestage,
         Watershed = watershed) %>%
  mutate(Lifestage = recode(Lifestage,
                            "juv_summer" = "Parr",
                            "redds" = "Redds",
                            "juv_winter" = "Presmolts")) %>%
  mutate(Lifestage = factor(Lifestage,
                            levels = c("Redds",
                                       "Parr",
                                       "Presmolts"))) %>%
  mutate(tot_cap_cv = tot_cap_se / tot_cap,
         Capacity = paste0(prettyNum(round(tot_cap), big.mark = ","), " (", round(tot_cap_cv, 2), ")")) %>%
  select(Lifestage:source, tot_length, Capacity) %>%
  rename(Length = tot_length) %>%
  pivot_wider(names_from = "source",
              values_from = c("Length", "Capacity"),
              names_glue = "{source}_{.value}") %>%
  kable(digits = 1,
        caption = "Estimated capacities (CVs) for Chinook in various lifestages using the master sample points and linear network reaches for extrapolation. Total watershed length for each method is shown as well.") %>%
  kable_styling(fixed_thead = T)
```


## Steelhead

```{r sthd-parr-fig, fig.cap = "Various scenarios of steelhead parr abundance including QRF capacity estimates."}
bar_list_p$Steelhead.Parr
```

```{r sthd-winter-fig, fig.cap = "Various scenarios of steelhead presmolt over-winter abundance including QRF capacity estimates."}
bar_list_p$Steelhead.Presmolts
```

```{r sthd-redd-fig, fig.cap = "Various scenarios of steelhead redd abundance including QRF capacity estimates."}
bar_list_p$Steelhead.Redds
```

```{r sthd-cap-tab}
cap_comp_df %>%
  filter(Species == "Steelhead") %>%
  rename(Lifestage = lifestage,
         Watershed = watershed) %>%
  mutate(Lifestage = recode(Lifestage,
                            "juv_summer" = "Parr",
                            "redds" = "Redds",
                            "juv_winter" = "Presmolts")) %>%
  mutate(Lifestage = factor(Lifestage,
                            levels = c("Redds",
                                       "Parr",
                                       "Presmolts"))) %>%
  mutate(tot_cap_cv = tot_cap_se / tot_cap,
         Capacity = paste0(prettyNum(round(tot_cap), big.mark = ","), " (", round(tot_cap_cv, 2), ")")) %>%
  select(Lifestage:source, tot_length, Capacity) %>%
  rename(Length = tot_length) %>%
  pivot_wider(names_from = "source",
              values_from = c("Length", "Capacity"),
              names_glue = "{source}_{.value}") %>%
  kable(digits = 1,
        caption = "Estimated capacities (CVs) for steelhead in various lifestages using the master sample points and linear network reaches for extrapolation. Total watershed length for each method is shown as well.") %>%
  kable_styling(fixed_thead = T)
```

## Lemhi Specific

We broke down the Lemhi available capacity into the Upper and Lower Lemhi (defined by the confluence of Hayden Creek).

```{r}
upper_lem_poly = upper_lem$basin
lower_lem_poly = cap_est %>%
  filter(watershed == "Lemhi") %>%
  slice(1) %>%
  pull(inputs) %>%
  map("wtsd_poly") %>%
  extract2(1) %>%
  extract2(1) %>%
  st_difference(upper_lem_poly)

lem_cap = cap_est %>%
  filter(watershed == "Lemhi") %>%
  select(lifestage:source) %>%
  mutate(upp_lem_cap = map(inputs,
                           .f = function(y) {
                             calc_watershed_cap(wtsd_polygon = upper_lem_poly,
                                                capacity_sf = y$rch[[1]],
                                                spp_range = y$spp_domain[[1]],
                                                by_stream = F) %>%
                               rename(n = n_rchs)
                           })) %>%
  mutate(low_lem_cap = map(inputs,
                           .f = function(y) {
                             calc_watershed_cap(wtsd_polygon = lower_lem_poly,
                                                capacity_sf = y$rch[[1]],
                                                spp_range = y$spp_domain[[1]],
                                                by_stream = F) %>%
                               rename(n = n_rchs)
                           })) %>%
  pivot_longer(cols = ends_with("lem_cap"),
               names_to = "area",
               values_to = "cap") %>%
  unnest(cols = cap) %>%
  mutate(area = recode(area,
                       "upp_lem_cap" = "Upper Lemhi",
                       "low_lem_cap" = "Lower Lemhi")) %>%
  select(-inputs) %>%
  mutate(across(tot_length,
                ~. / 1000))

lem_comp_df = lem_cap %>%
  rename(Watershed = watershed,
         Lifestage = lifestage) %>%
  mutate(Lifestage = recode(Lifestage,
                            "juv_summer" = "Parr",
                            "redds" = "Redds",
                            "juv_winter" = "Presmolts")) %>%
  mutate(Scenario = paste0("QRF_cap_", source)) %>%
  select(Lifestage:area,
         Scenario,
         Abundance = tot_cap) %>%
  filter(source == "Points") %>%
  bind_rows(rec_goals %>% 
              filter(Watershed == "Lemhi",
                     Lifestage %in% c("Parr", "Redds", "Presmolts"))) %>%
  mutate(Lifestage = factor(Lifestage,
                            levels = c("Redds",
                                       "Parr",
                                       "Presmolts")),
         Scenario = factor(Scenario,
                           levels = c("QRF_cap_Points",
                                      "Mean",
                                      "Max",
                                      "MAT",
                                      "MAT+25%"))) %>%
  mutate(area = if_else(is.na(area), as.character(Scenario), area)) %>%
  filter(Scenario %in% c("QRF_cap_Points",
                         "MAT+25%")) %>%
  mutate(Scenario = fct_expand(Scenario, "Available Capacity (QRF)", "Required Capacity\n(MAT +25%)"),
         Scenario = fct_recode(Scenario,
                               "Available Capacity (QRF)" = "QRF_cap_Points",
                               "Required Capacity\n(MAT +25%)" = "MAT+25%"),
         Scenario = fct_rev(Scenario)) %>%
  mutate(area = factor(area,
                       levels = c("MAT+25%",
                                  "Upper Lemhi",
                                  "Lower Lemhi"))) %>%
  arrange(Species, Lifestage, area, Scenario)

lem_bar_p = lem_comp_df %>%
  split(list(.$Species, .$Lifestage)) %>%
  map(.f = function(df) {
    df %>%
      ggplot(aes(y = Scenario,
                 x = Abundance,
                 fill = area)) +
      geom_col(position = "stack") +
      scale_fill_manual(values = c("Lower Lemhi" = "darkblue",
                                   "Upper Lemhi" = "lightblue",
                                   "MAT+25%" = "black"),
                        breaks = c("MAT+25%", "Lower Lemhi", "Upper Lemhi"),
                        name = NULL) +
      # scale_fill_brewer(palette = "Paired",
      #                   name = "Scenario") +
      theme(legend.position = "bottom",
            # axis.text.y = element_blank(),
            axis.title.y = element_blank(),
            axis.ticks.y = element_blank()) +
      scale_x_continuous(labels = function(x) format(x, scientific = F)) +
      labs(x = "Abundance",
           title = paste(unique(df$Species), unique(df$Lifestage)))
  })

  
```

### Chinook

```{r lem-parr-chnk-fig, fig.height = 2, fig.cap = "Overall recovery goal of Chinook parr in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Chinook.Parr
```

```{r lem-winter-chnk-fig, fig.height = 2, fig.cap = "Overall recovery goal of Chinook presmolts (overwintering) in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Chinook.Presmolts
```

```{r lem-redd-chnk-fig, fig.height = 2, fig.cap = "Overall recovery goal of Chinook redds in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Chinook.Redds
```

### Steelhead

```{r lem-parr-sthd-fig, fig.height = 2, fig.cap = "Overall recovery goal of steelhead parr in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Steelhead.Parr
```

```{r lem-winter-sthd-fig, fig.height = 2, fig.cap = "Overall recovery goal of steelhead presmolts (overwintering) in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Steelhead.Presmolts
```

```{r lem-redd-sthd-fig, fig.height = 2, fig.cap = "Overall recovery goal of steelhead redds in the Lemhi (MAT +25%) along with available capacity, colored by lower and upper Lemhi."}
lem_bar_p$Steelhead.Redds
```

```{r save-plots, eval = F}
walk(names(lem_bar_p), function(.x) {
  ggsave(
    path = "figures/",
    filename = paste0(.x, ".png"),
    plot = lem_bar_p[[.x]],
    width = 6.5,
    height = 2
  )
})
```

